---
title: Visualising Google location data in R
subtitle: An example using honeymoon loacation data in New Zealand
output: github_document
---

Having a go at visualising my Google location history data with R. It's not perfect.

## Required packages
```{r, warning=FALSE, message=FALSE}
library(rjson) # The google location history file is a JSON 
library(ggmap) # For ggplot2 background layer maps
library(tidyverse) # For ease of data wrangling
library(lubridate) # For date manipulation
```

## Data in
Read the raw json
```{r}
raw <- fromJSON(file='Takeout/Location History/Location History.json')
```

## Data wrangling
Extract the fields I require from the raw JSON.
```{r}
# Flatten the raw JSON file
flat <- flatten(flatten(raw))

# Extract the timestamp, lat, long and accuracy data
ts <- flat[names(flat) == "timestampMs"] %>% unlist() %>% unname()
lat <- flat[names(flat) == "latitudeE7"] %>% unlist() %>% unname()
lon <- flat[names(flat) == "longitudeE7"] %>% unlist() %>% unname()
acc <- flat[names(flat) == "accuracy"] %>% unlist() %>% unname()
```

Create a dataframe containing all records and add new variables. 
```{r}
df <- 
  tibble(ts=ts, lat=lat, lon=lon, acc=acc) %>%
  mutate(myts = as_datetime(as.numeric(ts)/1000)) %>% # Convert raw timestamp to datetime
  arrange(myts) %>% 
  mutate(year = year(myts),
         month = month(myts, label=T),
         lat = lat / 1e7, # I think these need to be divided by 1e7
         lon = lon / 1e7, # I think these need to be divided by 1e7
         xstart = lon, # Create start and end columns for plotting segments later on
         xend = lead(lon), 
         ystart = lat, 
         yend = lead(lat),
         dist = (((xend - xstart)^2) + ((yend - ystart)^2))^0.5) # Create a distance column (pythagoras)
```

## Get map data
Get coordinates of the box that surrounds New Zealand. I found this website http://bboxfinder.com which was useful for this.
```{r}
nz_box <- c(left= 166, right=179, bottom=-47.5, top=-34)
```

Get map using ggmap(). Here I have saved the map locally and load it in the building of this markdown document - hence the commented out lines.
```{r}
# nz_map <- get_stamenmap(bbox = nz_box, maptype = "toner-lite", zoom=7)
# save(nz_map, file="nz_map_stamen_tonerlite_zoom7.Rdata")
load(file="nz_map_stamen_tonerlite_zoom7.Rdata")
```

### Reduce data
As the raw JSON contains all of my location history data, here I reduce it to just records within the New Zealand bounding box. I also filter to records where the distance between points is less than 5 - this means segment lines are not drawn between points that are very far apart (such as the internal flight from Christcjurch to Auckland)
```{r}
nz_df <- 
  df %>%
  filter(between(lon, nz_box[1], nz_box[2]),
         between(lat, nz_box[3], nz_box[4])) %>% 
  filter(dist < 5)
```

## Visualise
```{r, fig.width=6, fig.height=8, fig.align='center'}
ggmap(nz_map) + 
  geom_rect(aes(xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf), fill="white", alpha=0.2)+
  geom_point(data = nz_df, aes(lon, lat, col=as.numeric(myts), size=acc))+
  geom_segment(data = nz_df, aes(x=xstart, xend=xend, y=ystart, yend=yend))+
  scale_color_viridis_c(option = "plasma", end=0.9, alpha=1,
                        breaks=as.numeric(range(nz_df$myts)),
                        labels=format(range(nz_df$myts), "%d %b %Y"))+
  scale_size(range=c(2,6), guide = FALSE)+
  theme_bw()+
  theme(legend.position = "right")+
  labs(x = "Longitude",
       y = "Latitude",
       title = "Honeymoon down under",
       subtitle = "Google location tracking",
       colour = "")
```

### Attempting density visualisation
I want to raster a 2D density estimate of the location history. I've done something similat using `geom_density_2d()`, but I want to be able remove density tiles below a certain threshold, so the density estimate does not cover the whole map (only loaclised areas of high density) - so I here I gave a go at producing that view.
```{r}
# Create the 2D density estimate
de2d <- MASS::kde2d(nz_df$lon, nz_df$lat, h=c(0.3,0.3), n=200, lims=nz_box)

# Wrangle the density estimate into a dataframe for use with geom_tile()
dendf <- 
  crossing(lat = de2d$y, lon = de2d$x) %>% 
  mutate(z = as.vector(de2d$z),
         z = z - min(z),
         z = z/max(z)) %>%  
  filter(z >= 0.0000001) # Only keep tiles above a certain density threshold
```

Visulise the plot with a `geom_tile()` layer
```{r, fig.width=6, fig.height=8, fig.align='center'}
ggmap(nz_map) + 
  geom_rect(aes(xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf), fill="white", alpha=0.2)+
  geom_tile(data = dendf, aes(lon, lat, fill=z), alpha=0.75)+
  scale_fill_viridis_c(option = "plasma", guide=FALSE)+
  theme_bw()+
  labs(x = "Longitude",
       y = "Latitude",
       title = "Honeymoon down under",
       subtitle = "Google location tracking",
       fill = "")
```

